---
title: "Functions and plot_ly figures from Flight Data"
description: |
  The following is an assignment I completed which includes sorting a data set to find the desired data, creating functions which can then be reused later, and finally creating figures to represent the data.
author:
  - name: Morelle Tinsley
    url: https://www.google.com
date: 04-20-2021
output:
  distill::distill_article:
    self_contained: false
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

```{r}
knitr::opts_chunk$set(echo = TRUE)

#first load the libraries we'll need
library(tidyverse)
library(plotly)
library(nycflights13) # <- make sure to install this package 


```

## Functions

1.  Create a function that will return the sum of 2 numbers.

```{r}
two_two <- function(firstnum, secondnum){
  
  result <- sum(firstnum, secondnum) 
  
  return(result)
  }
two_two(2,2) #ran the code with two numbers



```

2.  Create a function that will return the mean of a vector of numbers fed to it.\
    -You can see how I first created a function which would calculate a mean, and then later I input a vector of numbers, and the function found its mean for me.

```{r}

mynumbers <- sample(1:100, 10)

mean_function <- function(vector){ 
  
  result <- mean(vector)
  
  return(result)
}

mean_vector <- c(7,19,245,300,50,89)


mean_function(mean_vector)


```

3.  Now, expand on the function you made above, and instead of just returning the number, have it return the sentence:\
    "The mean is \_\_\_" where the blank is the calculated result.

```{r}

mean_vector <- c(7,19,245,300,50,89)

question3 <- function(vector.set) {
  avg.set <- mean(vector.set)
  print(paste("The mean is", avg.set))
}

question3(mean_vector)


```

```{r}
#load the data to use for the rest of the assignment questions
flights <- nycflights13::flights

```

4.  Using the `flights` dataset from the package `nycflights13` loaded in the above cell, create a function that filters the data frame to only include records from a single originating airport (column name `origin`).

```{r}



head(flights)

snglorgn <- function(flight_numbers) {
  flight_numbers %>%
    filter(origin == "JFK")
}

snglorgn(flights)





```

5.  Similarly, using the `flights` dataset, create a function that:\

-   filters the data frame to only include records from a single originating airport (column name `origin`), and then\
-   aggregates the results by airline (`carrier`) to show how many flights each airline has from that airport.

```{r}

head(flights)

snglorgn <- function(flight_numbers) {
  flight_numbers %>%
    filter(origin == "JFK") %>% 
    count(carrier)
}


snglorgn(flights)





```

6.  Similarly, using the `flights` dataset, create a function that:\

-   filters the data frame to only include records from a single originating airport (column name `origin`), and then\
-   filter to only include flights longer than a specific number of minutes (`air_time`), and then\
-   aggregates the results by airline (`carrier`) to show how many flights of that length each airline has from that airport.

```{r}

head(flights)

snglorgn_plusmore <- function(flight_numbers) {
  flight_numbers %>%
    filter(origin == "JFK") %>% 
    filter(air_time>400) %>% 
    count(carrier)
}

snglorgn_plusmore(flights)


```

## Plotly charts

7.  Using the `flights` dataset, use plotly to create a bar chart showing the breakdown of flights by airline that originate from JFK airport (`origin`) and are more than 400 minutes in length (`air_time`).

    -   Now we have come to the portion of the assignment where we use sorting and functions to make graphics to show our data.

    -   Below, you can see that I created a function which sorted my data, gave it a name, and then imput it into the Plot_ly chart which makes a bar chart from the newly filtered data.

```{r}
head(flights)

snglorgn_plusmore <- function(flight_numbers) { #made a function to filter the data
  flight_numbers %>%
    filter(origin == "JFK") %>% 
    filter(air_time>400) %>% 
    count(carrier)}

snglorgn_plusmore(flights)
  
JFK_bar <- snglorgn_plusmore(flights)

plot_ly(data = JFK_bar,                    #create a bar chart from it
        x = ~carrier,
        y = ~n,
        type = "bar"
        ) %>% 
  layout(title = "400 minute flights by carrier from JFK",
         xaxis = list(title = "Carrier"),
         yaxis = list (title = "Number of flights"))





```

8.  Using the `flights` dataset, use plotly to create a bar chart showing which destination airports (`dest`) have the most flights with durations of over 400 minutes (`air_time`).

```{r}

head(flights)

dest_count <- function(flight_numbers) {
  flight_numbers %>%
    filter(air_time>400) %>% 
    count(dest)}

dest_count(flights)
  
Dest_bar <- dest_count(flights)

plot_ly(data = Dest_bar,
        x =~dest,
        y = ~n,
        type = "bar") %>% 
  layout(title = "Destinations with most 400 minute flights",
         xaxis = list(title = "Destination"),
         yaxis = list (title = "Number of flights"))




```

9.  Using the `flights` dataset, use plotly to create a donut chart showing the share of flights each airline has for flights with a destination of Daniel K. Inouye International Airport (HNL) in Honolulu.

```{r}
HNL_share <- function(flight_numbers){
  flight_numbers %>% 
    count(dest == "HNL")
}
```

```{r}

head(flights)

flights %>% 
  filter(dest =="HNL") %>% 
  count(carrier) 
```

```{r}

```

```{r}
carrier_count <- function(data){
results <- data %>% 
    filter(dest == "HNL") %>% 
  count(carrier) 

return(results)
}
```

```{r}
honolulu_donut <- carrier_count(flights)
```

```{r}


fig <- honolulu_donut %>% plot_ly(labels = ~carrier, values = ~n)

fig <- fig %>% add_pie(hole = 0.6) #would make a pie chart, to make a donut chart add the hole

fig <- fig %>% layout(
                  title = "Number of HNL flights by carrier",  showlegend = F,
                  xaxis = list(showgrid = FALSE, zeroline = FALSE, showticklabels = FALSE),
                  yaxis = list(showgrid = FALSE, zeroline = FALSE, showticklabels = FALSE))
fig

```

10. Using the `flights` dataset, use plotly to create a scatterplot chart for flights with a destination of Daniel K. Inouye International Airport (HNL) in Honolulu, that shows the relationship between the departure delay (`dep_delay`) and arrival delay (`arr_delay`).

```{r}

head(flights)

delay_data <- flights %>% 
  filter(dest == "HNL")
```

```{r}
delay_plot_ly<- plot_ly(data = delay_data, x = ~dep_delay, y = ~arr_delay)

delay_scatter <- plot_ly(data = delay_data, x = ~dep_delay, y = ~arr_delay) %>% 
  layout(title = "Honolulu Arrival Delays According to Departure Delays")

delay_scatter
```

\`\`\`
