---
title: "Mapping in Leaflet"
description: |
  Here is a sample of maps I made using Leaflet to represent different data sets in R.
author:
  - name: Morelle Tinsley
    url: https://www.google.com
date: 04-20-2021
output:
  distill::distill_article:
    self_contained: false

---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)


# Load packages
library(tidyverse)
library(lubridate)
library(readxl)
library(janitor)
library(glue)
library(leaflet)
library(RColorBrewer)
library(htmlwidgets)
library(htmltools)
library(tigris)
library(tmap)
library(tmaptools)
options(tigris_class = "sf")
```

### Map of COVID-19 Deaths by State as of February 22, 2021

First, I loaded in COVID-19 data from the New York Times. I filtered this data to show the specified date, which is the day the US surpassed 1million deaths due to COVID-19. Then I input this filtered data into a leaflet map, and added colors and the interactive stats which show up when you hover over a specific state. Finally, I added a map legend.

```{r include=FALSE}
covidcases <- read_csv("https://raw.githubusercontent.com/nytimes/covid-19-data/master/us-states.csv")
```

```{r include=FALSE}
#Download the polygon boundaries for states at the lowest resolution using tigris package
states <- tigris::states(cb=T)
```

```{r include=FALSE}
covid_data <- covidcases %>% 
  filter(date == "2021-02-22") #filter and name the data set

covid_data

covid_merged <- geo_join(states, covid_data, "NAME", "state") #merge states with covid data
```

```{r}
leaflet() %>%
  addTiles() %>%
  setView(-98.483330, 38.712046, zoom = 4) 

mypalette <- colorNumeric("YlOrRd", domain = covid_merged$deaths) #create a color palette for the map

popup_cd <- paste0("Total Deaths: ", as.character(covid_merged$deaths)) #make the hover over thingy

covid_death_map<-leaflet() %>%
  addTiles() %>%
  setView(-98.483330, 38.712046, zoom = 4) %>%   #set the view and zoom to show what you want to see (in our case the continental US)
  addPolygons(data = covid_merged , 
              fillColor = ~mypalette(covid_merged$deaths), #fill the states by darkening color=more deaths
              fillOpacity = 0.9, 
              weight = 0.2, 
              smoothFactor = 0.2,
              label=popup_cd, #add the hover over total deaths labels
              labelOptions = labelOptions(
                style = list("font-weight" = "normal", padding = "3px 8px"),
                textsize = "15px",
                direction = "auto")) %>%
  addLegend(pal = mypalette, #add a map legend
            values = covid_merged$deaths,
            bins = 4,
            position = "bottomright", 
            title = "Covid Cases", 
            labFormat = labelFormat(digits = 0))

covid_death_map
```

### Map of Sacramento Real Estate Sales

The following map expresses data of real estate sales in Sacramento. Using leaflet, I added circles whose size corresponds to the price of the real estate sale in that particular latitude and longitude location.

```{r}

sac_sales <- read_csv("https://support.spatialkey.com/wp-content/uploads/2021/02/Sacramentorealestatetransactions.csv")

sac_sales <- sac_sales %>% 
  filter(type == "Residential") #filter out the rows we don't want

sac_map <- leaflet(sac_sales) %>% #create a map in leaflet
  addTiles() %>%
  setView(-121.4944, 38.5816, zoom = 11) %>% #set the coordinates of sacramento and zoom accordingly
  # addTiles() %>%
  addProviderTiles(providers$Stamen.TonerLite) %>% #alternative tile background
  addCircles(lng = ~longitude, #add the circles on the coordinents of the sales
             lat = ~latitude, 
             weight = 1,
             # stroke = FALSE,
             # fillOpacity = .50,
             radius = ~sqrt(price) * 1 #set the size of the circles relative to the price of the sale
  ) %>%
  addControl("Real Estate Sales (itemized) by zip code", position = "topright") #add title

sac_map
```
